### 2019-08-20 maciek.tex
Lower bound for augmentation

\begin{theorem}
  The competitive ratio of any deterministic algorithm for Online Balanced Partition (with existance of a perfect partition of the input graph) is $\Omega(\ell)$.
\end{theorem}
%Note that in the scenario with a perfect partition, both OPT and ALG always pays at most $O(\ell)$.

@@

Consider clusters $A$ and $B$ that contain nodes $a_1, a_2, a_3$ and $b_1, b_2, b_3$, respectively.
The adversary begins with issuing three requests: $(a_1, a_2)$, $(b_1, b_2)$, and $(a_3, b_3)$.
Note that these require moving one of these pairs to a third cluster.
Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a carefuly chosen cluster, that we denote by $D$), and for OPT all other requests are free.

The algorithm must place these three components in different clusters. Let's call the cluster of $(a_1, a_2)$ the cluster $A_{ALG}$, the cluster of $(b_1, b_2)$ the cluster $B_{ALG}$ and the cluster of $(a_3, b_3)$ the cluster $C$. The configuration of ALG's clusters is then: $A_{ALG}$ contains $a_1, a_2,x$, the cluster $B_{ALG}$ contains $b_1, b_2,y$ and the cluster $C$ contains $a_3, b_3, z$, for some $x, y, z$ (of the algorithm's choice).
For this, ALG pays at least $2$.

The OPT places these three components in clusters $A, B$ and $D$, paying $2$ for migrating $a_3$ and $b_3$. The OPT has the freedom of choice of the cluster $D$ depending on the choices of the algorithm. We choose $D$ in such way that it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$'s or $y$'s place as the last possible move \maciek{elaborate}.
The configuration of OPT's clusters is then: $A$ contains $a_1, a_2,d_1$, $B$ contains $b_1, b_2,d_2$ and $D$ contains $a_3, b_3, d_3$.
Note that the requests between nodes of the same color are free for OPT, with the exception of nodes from $A$, $B$, and most notably $D$.

@@

>>>>

\begin{theorem}
  The competitive ratio of any deterministic algorithm for Online Balanced Partition (assuming the existence of a perfect partition of the input graph) is $\Omega(\ell)$.
  \label{th:lb_omega_l}
\end{theorem}
%Note that in the scenario with a perfect partition, both OPT and ALG always pays at most $O(\ell)$.

@@

Consider clusters $A$ and $B$ containing nodes $a_1, a_2, a_3$ and $b_1, b_2, b_3$, respectively.
The adversary begins with issuing three requests: $(a_1, a_2)$, $(b_1, b_2)$, and $(a_3, b_3)$.
Note that the algorithm is forced to move at least one of these components to a third cluster:
in general, if the algorithm does not keep each component within one cluster, the adversary might continue issuing requests among nodes of this component.

The algorithm places these three components in different clusters. Let's call the cluster of $(a_1, a_2)$ the cluster $A_{ALG}$, the cluster of $(b_1, b_2)$ the cluster $B_{ALG}$ and the cluster of $(a_3, b_3)$ the cluster $C$. The configuration of ALG's clusters is then: $A_{ALG}$ contains $a_1, a_2,x$, the cluster $B_{ALG}$ contains $b_1, b_2,y$ and the cluster $C$ contains $a_3, b_3, z$, for some $x, y, z$ (of the algorithm's choice).
For this, ALG pays at least $2$.

Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a carefuly chosen cluster, that we denote by $D$), and for OPT all other requests are free. 
The OPT places these three components in clusters $A, B$ and $D$, paying $2$ for migrating $a_3$ and $b_3$. The OPT has the freedom of choice of the cluster $D$ depending on the choices of the algorithm. We choose $D$ in such way that it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$'s or $y$'s place as the last possible move \maciek{elaborate}.
The configuration of OPT's clusters is then: $A$ contains $a_1, a_2,d_1$, $B$ contains $b_1, b_2,d_2$ and $D$ contains $a_3, b_3, d_3$.
Note that the requests between nodes of the same color are free for OPT, with the exception of nodes from $A$, $B$, and most notably $D$.

@@

\begin{observation}
  The lower bound of $\Omega(\ell)$ can be generalized for arbitrary $k = c\cdot 3$.
  The adversary starts with a serie of ``glueing requests'', so that each cluster contains 3 components of size $k/3$.
  We repeat the argument of Theorem~\ref{th:lb_omega_l} using these glued components instead of singletons.
  OPT pays $O(k)$ for migration of one such component, and ALG pays $\Omega(k \cdot \ell)$.
  \label{obs:lb_general_k}
\end{observation}

\begin{observation}
  The lower bound of $\Omega(\ell)$ holds even with $1/3-1/k$-augmentation, i.e., in scenarios where the capacity of each cluster of the algorithm is $(1+1/3)\cdot k - 1$, while the capacity of OPT's clusters is $k$.
  To this end we use the construction from Observation~\ref{obs:lb_general_k}.
  We observe that each cluster of the algorithm can fit exactly $3$ components of size $k/3$.
  \maciek{Note that this does not contradict the existence of $O((1+1/\epsilon)\cdot k \cdot \log(k))$ algorithm of Avin, Bienkowski et al, as it requires $2+\epsilon$-augmentation.}
\end{observation}
